\Large Co-evolution of lexical meaning \& pragmatic use\\ {\bf Revision Cover letter}

### 2018-03-19 paper/rv-letter1.tex
repsonse letter

\vspace{.5cm}
\begin{mdframed}[backgroundcolor=gray!25,linecolor=gray!25]
Does the model assume that learners observe not only the behaviour (the message, e.g. agent A hearing agent B say: "I own some of JC's albums") but also the true state of the world (i.e. agent A also knows whether agent B owns all or some-but-not-all albums)? Or do they get feedback on their communicative success? In other words, how do they learn the correct state-message mappings that form the lexicon? 
\end{mdframed}
Learners witness sequences of data $d$. Such a sequence is composed of message-state
pairings. In other words, learners do witness the true state of the world (Section~2.3.3, {\em
  learnability}). This, however, is a widely made assumption in the literature on iterated
learning.

@@

We do not believe that, ultimately, the non-parsimonious view that there are different complexity measures for different structures should be assumed. However, with previous work in the iterated learning tradition, we do assume that (i) inductive learning biases can play a role in shaping culturally transmitted knowledge and that, as the reviewer notes, (ii) biases can be of manifold nature; another well-studied example being the mutual exclusivity bias, which could be expected to work against targets and in favor of competitors.

At present, it is impossible to disentangle the complex interaction of multiple biases. We therefore focus on the effects of a single plausible contributing factor to a preference of target lexica over competitors: a well-motivated bias favoring simplicity \citep{feldman:2000,chater+vitanyi:2003, piantadosi+etal:2012a,kirby+etal:2015,piantadosi+etal:underreview}. By assumption, this translates to a preference not to lexicalize linguistic material over lexicalizing it; simpler formulae over complex ones. As mentioned in Section 3.1.2, the complexity measure that the prior is based on is just a convenient operationalization of this particular hypothesis.

We should stress that, just as models of iterated learning make different predictions under the assumption of different inductive biases, so does our model. The assumption and operationalization of a particular bias should, of course, always be seen critically (Section 4).

@@

\begin{mdframed}[backgroundcolor=gray!25,linecolor=gray!25,frametitle= Reviewer \thereviewerCounter~comment \thereviewerCommentCounter \hfill ~~({\it transmission fidelity})]
%
What is your measure of fidelity/infidelity? Is it just 0 if the teacher and learner are of a different type and 1 if they are of the same? Could you have a more graded measure of fidelity, e.g. taking into account if two types are of the same 'kind' or not? 
\end{mdframed}
%
We understand transmission fidelity as the probability of acquiring type $i$ when learning from type $i$ (Section~2.3.3). This is the value of a cell in the learning matrix $Q$. In this case, $Q_{ii}$. As discussed in Section~2.2, if $Q_{ii} = 1$ then $i$ is always acquired when learning from type $i$. On the other extreme, $Q_{ii} = 0$ means that $i$ is never acquired from $i$. $Q$ is a stochastic matrix and we we usually see neither extreme but rather, as suggested by the reviewer, a graded notion of transmission fidelity (Section~2.2). The details on how $Q$ is computed are given in Section 2.3.3. We have expanded this section to make these details clearer in the prose.
%

@@

We have thoroughly revised the manuscript to make the contrast between these three predictions clearer (see Section~3.2, particularly 3.2.3; but also Section~2.4, which illustrates how the dynamics behave in a smaller type space). These questions tie in with Reviewer $1$'s request for a clearer exposition and the explanation of why we speak of co-evolution in this case (reviewer 1, comment 1). In a nutshell, pressure for learnability alone does not put the types in a population in competition. Instead, the population comes to be inhabited by types that are recovered from learning input of replicated types more often -- of which there might be many (e.g., all variants of a kind are equally learnable; if there are many variants of a type that is, in tendency, acquired more often, then the population will stay polymorphic). What fitness-relative selection adds to this process is the competition that learnability alone lacks, independent of the type space we look at. Now, in the case of our type space we have a slight (modulo $\lambda$) communicative disadvantage for targets relative to competitors, for instance. However, the fact that targets are inferred more more often by na\"ive learners, also from those that learn from competitors, leads (i) to the gradual existence of less competitors, counteracting this functional disadvantage, as well as (ii) to the existence of only one variant of a kind (contra what we get in the {\em learning alone} condition).
%

@@

As noted above, we have expanded our analysis of the model and put much more emphasis on how the components that feed it play a role in driving particular evolutionary outcomes. See, in particular, Section 2.4, Section 3.2, and Section 4.
%

@@

It is not necessarily true that it will always be advantageous for agents to behave pragmatically (see the end of Section~2.3.1 but also the discussion surrounding Figure~3a).  The main thing to note is that a pragmatic level-$1$ hearer using lexicon $L_{\text{bound}}$ will reason about the behavior of a {\em soft-maximizing} level-$0$ speaker of $L_{\text{bound}}$. The lower $\lambda$, the more the stochasticity percolates from the level-$0$ speaker to level-$1$ pragmatic interpretation. By contrast, literal level-$0$ hearers of $L_{\text{bound}}$ have a one-to-one form-meaning mapping from the get go. Put differently, pragmatic reasoning can actually encumber some hearers, depending on their lexicon. Particularly if $\lambda$ is low.

@@

As we now clarify in Section~2.3.1 and stress a couple of times throughout the article, we assume very little sophistication from our agents. In particular, they do not know the type of the player that the are playing against. They simply behave according to their subjective point of view in a boundedly rational fashion. For example, the linguistic choices of a level-$1$ speaker of lexicon $L_{\text{some}}$ do not change depending on whom she interacts with. They are always defined as (boundedly) rational choice relative to the interpretative behavior of a level-$0$ user of $L_{\text{some}}$ (see definitions (3) --(6) in \S2.3.1).
%

@@

\begin{mdframed}[backgroundcolor=gray!25,linecolor=gray!25,frametitle= Reviewer \thereviewerCounter~comment \thereviewerCommentCounter \hfill ~~({\it posterior type sampling})]
%
The learner in this model always selects a particular grammar/pragmatic type. An alternative would be for them to maintain uncertainty about the type (which is optimal behavior from a Bayesian perspective). I suspect that this model is intractable (or at least much less manageable) than the proposed one, but it would be worth noting that this is a substantive choice point in the model.
\end{mdframed}
We now highlight that this is a design choice and that there are alternatives in Section~2.3.3.

@@

@@

This is an important point. We agree that some fragments of our previous manuscript could lead to confusion, e.g., as to why $L_{\text{all}}$, being the {\em a priori} more likely type to be inferred, did not come out strongly even in the {\em learning only} condition. Throughout the article, we now stress more that what matters is the posterior, and put emphasis on the role of the likelihood.

@@

As now made explicit in Section~3.1, we chose three messages for illustrative purposes mainly. In particular, we wanted to inspect a type space where targets are not the most likely {\em a priori} as this may mislead readers into thinking that learnability alone, in particular the prior alone, drives the outcome. A larger type space also better allows us to showcase that multiple kinds of a type may exist and how this affects evolutionary outcomes. The reason why we did not go beyond $3$ messages is for computational tractability. Calculating $Q$, even if approximated, is expensive. The question whether increasing the type space will decrease the frequency to which we expect to see pragmatic {\em some} users is interesting. At present, we have no answer.

>>>>

\vspace{.5cm}
\begin{mdframed}[backgroundcolor=gray!25,linecolor=gray!25]
Does the model assume that learners observe not only the behaviour (the message, e.g. agent A hearing agent B say: "I own some of JC's albums") but also the true state of the world (i.e. agent A also knows whether agent B owns all or some-but-not-all albums)? Or do they get feedback on their communicative success? In other words, how do they learn the correct state-message mappings that form the lexicon? 
\end{mdframed}
Learners witness sequences of data $d$. Such a sequence is composed of message-state
pairings. In other words, learners do witness the true state of the world (Section~2.4.3, {\em
  learnability}). This, however, is a widely made assumption in the literature on iterated
learning.

@@

First of all, the model does not rest on this assumption at all; only the particular application does. Following previous work in the iterated learning tradition, we do assume that (i) inductive learning biases can play a role in shaping culturally transmitted knowledge and that, as the reviewer notes, (ii) biases can be of manifold nature.

On the other hand, it is true that the model itself poses constraints on what might count as a plausible inductive bias. One well-studied example of an inductive bias from the iterated learning tradition is the mutual exclusivity bias. This bias should be expected to work against targets and in favor of competitors, as the reviewer says. Yet in the context of our model a bias for mutual exclusivity might be hard to motivate, because the model itself has an explicit component of fitness-based selection for communicative efficiency. This component should take the part of selecting for mutual exclusivity. In other words, the division of labor is shifted here: not every constraint is as easily justifiable; the model tries to give an explanation in terms or fitness-based selection for some of the effects that certain biases are meant to explain in systems which only look at iterated learning.

What counts for this model is chiefly a bias that is \emph{not} motivated by the effects it has on communicative efficiency. Instead we only focus on the effects of a single plausible contributing factor to a preference of target lexica over competitors: a well-motivated bias favoring representational simplicity of single concepts, not lexica as a whole \citep{feldman:2000,chater+vitanyi:2003, piantadosi+etal:2012a,kirby+etal:2015,piantadosi+etal:underreview}. How to spell out such a lexical simplicity bias is, we readily admit, an open question. As mentioned in Section 3.1.2, the complexity measure in terms of a Language of Thought is just a convenient operationalization of this particular hypothesis for the sake of a concrete working example.

We should stress that, just as models of iterated learning make different predictions under the assumption of different inductive biases, so does our model. The assumption and operationalization of a particular bias should, of course, always be seen critically. We criticize our own setup and assumptions in Section 4.

@@

\begin{mdframed}[backgroundcolor=gray!25,linecolor=gray!25,frametitle= Reviewer \thereviewerCounter~comment \thereviewerCommentCounter \hfill ~~({\it transmission fidelity})]
%
What is your measure of fidelity/infidelity? Is it just 0 if the teacher and learner are of a different type and 1 if they are of the same? Could you have a more graded measure of fidelity, e.g. taking into account if two types are of the same 'kind' or not? 
\end{mdframed}
%
We understand transmission fidelity as the probability of acquiring type $i$ when learning from type $i$ (Section~2.4.3). This is the value of a cell in the learning matrix $Q$. In this case, $Q_{ii}$. As discussed in Section~2.2, if $Q_{ii} = 1$ then $i$ is always acquired when learning from type $i$. On the other extreme, $Q_{ii} = 0$ means that $i$ is never acquired from $i$. $Q$ is a stochastic matrix and we we usually see neither extreme but rather, as suggested by the reviewer, a graded notion of transmission fidelity (Section~2.2). The details on how $Q$ is computed are given in Section 2.4.3. We have expanded this section to make these details clearer in the prose.
%

@@

We have thoroughly revised the manuscript to make the contrast between these three predictions clearer (see Section~3.3, particularly 3.3.3; but also Section~3.1, which illustrates how the dynamics behave in a smaller type space; as well as the discussion of iterated learning alone and its dependence on $\lambda$ and $l$ in Sections 2.4.3 and appendix C). These questions tie in with Reviewer $1$'s request for a clearer exposition and the explanation of why we speak of co-evolution in this case (reviewer 1, comment 1). In a nutshell, pressure for learnability alone does not put the types in a population in competition. Instead, the population comes to be inhabited by types that are recovered from learning input of replicated types more often -- of which there might be many (e.g., all variants of a kind are equally learnable; if there are many variants of a type that is, in tendency, acquired more often, then the population will stay polymorphic). What fitness-relative selection adds to this process is the competition that learnability alone lacks, independent of the type space we look at. Now, in the case of our type space we have a slight (modulo $\lambda$) communicative disadvantage for targets relative to competitors, for instance. However, the fact that targets are inferred more often by na\"ive learners (by asymmetries in production likelihoods, see appendix C; a tendency which is also mediated by $\lambda$ and $l$), leads (i) to the gradual existence of less competitors, counteracting this functional disadvantage, as well as (ii) to the existence of only one variant of a kind (contra what we get in the {\em learning alone} condition).
%

@@

As noted above, we have expanded our analysis of the model and put much more emphasis on how the components that feed it play a role in driving particular evolutionary outcomes. Section 3.1 can be seen a direct response to this comment, where we try to deduce from what is known from the literature as much insight as possible about the workings of the dynamic in this case. Unfortunately, owing to the complexity of the case study we are not able to offer a mathematical characterization of, say, the stable rest points of the dynamic. (Notice that for this purpose we should switch to the continuous time formulation, which would also extend the paper further.)
%

@@

It is not necessarily true that it will always be advantageous for agents to behave pragmatically (see the end of Section~2.4.1 but also the discussion surrounding Figure~3a).  The main thing to note is that a pragmatic level-$1$ hearer using lexicon $L_{\text{bound}}$ will reason about the behavior of a {\em soft-maximizing} level-$0$ speaker of $L_{\text{bound}}$. The lower $\lambda$, the more the stochasticity percolates from the level-$0$ speaker to level-$1$ pragmatic interpretation. By contrast, literal level-$0$ hearers of $L_{\text{bound}}$ have a one-to-one form-meaning mapping from the start. Put differently, pragmatic reasoning can actually encumber some hearers, depending on their lexicon. Particularly if $\lambda$ is low.

More to the point of the question, even if higher levels of pragmatic reasoning were always at least as good as the next lower level, the current population-level dynamic rests on \emph{minimal} assumptions about agent-level rationality. We have expanded on this point, which was definitely unclear in the first version, in section 2.3. In particular footnote 1 is almost a direct reply to this excellent comment. So, our agents are deliberately \emph{not} assumed to know the type of their interlocutor, but merely iterate pragmatic reasoning based on their own behavior. They would also not play a best response to the population average. The dynamic assumes that agents do not innovate strategies (like higher-order pragmatic reasoning). The dynamic is built on the idea that fitness-based selection is a non-innovative, gradient and proportional amplifier of whatever works well. Agents might never find out what works well, e.g., when fitness is just a measure of the proportion of offspring (a biological interpretation). Or they find out implicitly based on what works for them when they explore different ways of behavior (reinforcement learning). Or they occasionally realize that other agents are communicatively successful to a certain degree (e.g., conditional imitation based on success). Yet other interpretations are possible as well.

@@

Indeed, this was regrettably not clear enough in the previous version. As we now emphasize in Sections~2.3 and 2.4.1, we assume very little sophistication from our agents. In particular, they do not know the type of the player that the are playing against. They simply behave according to their subjective point of view in a boundedly rational fashion. For example, the linguistic choices of a level-$1$ speaker of lexicon $L_{\text{some}}$ do not change depending on whom she interacts with. They are always defined as (boundedly) rational choice relative to the interpretative behavior of a level-$0$ user of $L_{\text{some}}$ (see definitions (3) --(6) in \S2.3.1).
%

@@

\begin{mdframed}[backgroundcolor=gray!25,linecolor=gray!25,frametitle= Reviewer \thereviewerCounter~comment \thereviewerCommentCounter \hfill ~~({\it posterior type sampling})]
%
The learner in this model always selects a particular grammar/pragmatic type. An alternative would be for them to maintain uncertainty about the type (which is optimal behavior from a Bayesian perspective). I suspect that this model is intractable (or at least much less manageable) than the proposed one, but it would be worth noting that this is a substantive choice point in the model.
\end{mdframed}
We now highlight that this is a design choice and that there are alternatives in Section~2.4.3, footnote 3.

@@

More concretely, there is a technical reason why we would like all speaker types to produce any message with positive probability: we do not want to make types too easily identifiable by even small sets of observed data, so that iterated learning has some work to do. Moreover, we belief that while a logarithmic definition of speaker utilities is just fine, when it comes to evolutionary fitness, a definition in terms of beliefs is not. Fitness should track genuine observable interaction with the world. Just holding a ``good belief'' gives no countable or observable advantage unless it is acted upon in some manner.

@@

The agents in this model do not actually learn the ``grammar'' of concepts which is used only as a tool for fixing one concrete measure of complexity of lexical concepts. It is these lexical concepts that are learned. We have tried to make this clearer by more careful reformulation.

@@

As now made explicit in Section~3.2, we chose three messages for illustrative purposes mainly. In particular, we wanted to inspect a type space where targets are not the most likely {\em a priori} as this may mislead readers into thinking that learnability alone, in particular the prior alone, drives the outcome. A larger type space also better allows us to showcase that multiple kinds of a type may exist and how this affects evolutionary outcomes. The reason why we did not go beyond $3$ messages is for computational tractability. Calculating $Q$, even if approximated, is expensive. The question whether increasing the type space will decrease the frequency to which we expect to see pragmatic {\em some} users is interesting. At present, we have no answer.

The new section 3.1 now also looks at the case with just two messages that the reviewer brings into play. It is useful to contrast exactly why the effects of iterated learning hinge on the set of all types (the point made in connection with appendix C).
